### 2014-07-23 paper_ieee.tex

\documentclass[final,journal,letterpaper]{IEEEtran}

\usepackage{cite}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{balance}
\usepackage{float}
\usepackage{threeparttable}

\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{textcomp}
\usepackage{mathrsfs}
\usepackage{amsthm}

\usepackage{mathptmx}
\usepackage[scaled=.90]{helvet}
\usepackage{courier}

\usepackage{color}

\usepackage{listings}
\usepackage{listings}

\lstset{
   language=C,
   basicstyle=\small,
   keywordstyle=\bfseries,
   identifierstyle=\ttfamily,
   stringstyle=\ttfamily,
   numbers=left,
   numberstyle=\tiny,
   stepnumber=1,
   numbersep=-5pt,
   showstringspaces=false
%   frame=single %trbl%
}

\usepackage[normalem]{ulem}

%------------------------------------------------------------------------------

\newcommand{\etal}{\emph{et al.}}
\newcommand{\eg}{\emph{e.g.}}
\newcommand{\ie}{\emph{i.e.}}
\newcommand{\etc}{\emph{etc.}}

%\newcommand{\comment}[1]{\textcolor{red}{AA: \bf #1}}
\newcommand\comment[1]{} % When you enable this lines, all the notes will disappear

%\renewcommand{\baselinestretch}{0.98}\normalsize
%------------------------------------------------------------------------------
\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\theoremstyle{plain}
\newtheorem{thm}{\protect\theoremname}
\theoremstyle{definition}
\newtheorem{defn}{\protect\definitionname}
\theoremstyle{remark}
\newtheorem{rem}{\protect\remarkname}

%\newtheorem{defn}[thm]{\protect\definitionname}
%\theoremstyle{remark}
%\newtheorem{rem}[thm]{\protect\remarkname}
%\makeatother

%\usepackage{babel}
\providecommand{\definitionname}{Definition}
\providecommand{\remarkname}{Remark}
\providecommand{\theoremname}{Theorem}

\begin{document}

\author{Vincenzo~Catania, Andrea Araldo~and~Davide~Patti% 
\thanks{The authors are with the Dipartimento di Ingegneria
Elettrica, Elettronica ed Informatica, Universit\`a di Catania, Italy
(email: \{vincenzo.catania,aaraldo,davide.patti\}@dieei.unict.it).}}

\title{Parameter Space Representation of Pareto Front to
Explore Hardware-Software Dependencies}
%ATTENZIONE: in questo testo io mi sono riferito a ``Pareto set''
%come un sottoinsieme dello spazio degli obiettivi (vedi, ad es., descrizione
%dell'algoritmo). E' più corretto forse, invece, intendere il Pareto
%set come un sottoinsieme dello spazio delle configurazioni.

\maketitle
\begin{abstract}
Embedded Systems Design often requires conflicting objectives to be
optimized with an appropriate choice of hardware/software parameters .
Interdependencies among these parameters can make this task even more
difficult, since an a-priori knowledge of them would  be required to
develop an analytical model of the system. In this work we present PS,
a multiobjective exploration strategy that introduces the notion of
Parameters Space Representation. After a detailed formal description
of the algorithm and its concepts, we show a hardware/software
exploration case study. Qualitative and quantitative comparisons of PS
against a well-know multiobjective genetic approach demonstrate that,
while not outperforming it in terms of dominance, the proposed
approach can trade the uniformity and granularity qualities of the
solutions found for obtaining larger Pareto sets, thus representing a
further choice for designer when objective constrains allow it.
\end{abstract}

%------------------------------------------------------------------------------
%\IEEEkeywords{...}

\begin{IEEEkeywords}
DSE, MultiObjective Optimization, Genetic Algorithms
\end{IEEEkeywords}
%------------------------------------------------------------------------------
\input{introduction.tex}
\input{concepts.tex}
\input{algorithm.tex}
\input{experiments.tex}
%------------------------------------------------------------------------------

\section{Conclusions}
In this work we presented a multiobjective strategy which 
introduces the concept of Parameter Space representation of Pareto
Set. A case study of a VLIW architecture involving strong
hardware/software dependencies has been carried out to evaluate
effectiveness with different amount of simulations budget.
A qualitative/quantitative comparison against a widespread multiobjective genetic
approach showed how the proposed strategy can result in a different
distribution of Pareto points which trades solutions granularity with
improved pareto extension.

\balance

\bibliographystyle{IEEEtran} 
\bibliography{bibliography}

% TODO: G2012 bib not used ?

\end{document}