### 2014-07-24 concepts.tex

\section{Underlying concepts and theoretical formulation}
\comment{Attenzione: siamo capaci di avere a che fare con unfeasible configurations?}

\comment{Theoretical suona troppo pomposo? Esportare di nuovo tutte le immagini libreoffice - svg - eps che viene meglio e aggiutare i nomi degli assi}
In this section we provide the concepts and the formal definitions that we will use later to express our algorithm. \comment{Ha un nome l'algoritmo?} First, we give the definitions of Pareto-set and Pareto-front that lie on the basis of Design Space Exploration is based on. Then, we sketch the fundamental idea behind our algorithm. Finally, we provide a formal definition of the regions, of their properties and of the operations that will be performed on them by the algorithm.

\subsection{Pareto-set and Pareto-front}
\label{sec:pareto}

Let $S$ be a parameterized system with $n$ parameters. The generic
parameter $p_i, \ i \in \{1,2,\ldots,n\}$ can take any value in
the set $V_i$. A {\em configuration} $\mathbf{c}$ of the system
$S$ is a $n$-tuple $\langle v_1,v_2,\ldots,v_n \rangle$ in which
$v_i \in V_i$ is the value fixed for the parameter $p_i$. The {\em
configuration space} (or {\em design space}) of $S$ [which we will
indicate as $\mathcal{C}(S)$] is the complete range of possible
configurations [$\mathcal{C}(S) = V_1 \times V_2 \times \ldots
\times V_n$]. Naturally, not all the configurations of
$\mathcal{C}(S)$ can really be mapped on $S$, but only a subset 
is feasible. We call this subset {\em
feasible configuration space} $S$ and indicate it as
$\mathcal{C}^*(S)$].

Let $m$ be the number of objectives to be optimized (e.g. power,
cost, performance, etc.). An {\em evaluation function}
$E:\mathcal{C}^*(S)\times \mathcal{B} \longrightarrow \Re^m$ is a
function that associates each feasible configuration of $S$ with
an $m$-tuple of values corresponding to the objectives to be
optimized when any application belonging to the set of benchmarks
$\mathcal{B}$ is executed.

Given a system $S$, an application $b \in \mathcal{B}$ and two
configurations $\mathbf{c}', \mathbf{c}'' \in \mathcal{C}^*(S)$,
$\mathbf{c}'$ is said to {\em dominate} (or {\em eclipse})
$\mathbf{c}''$, if, given $\mathbf{o}'=E(\mathbf{c}', b)$ and
$\mathbf{o}''=E(\mathbf{c}'', b)$, it results that $\mathbf{o}'
\leq \mathbf{o}''$ and $\mathbf{o}' \neq \mathbf{o}''$, where
vector comparisons are interpreted component-wise and are true
only if all of the individual comparisons are true ($o'_i \leq
o''_i \ \forall \ i = 1,2,\ldots,m$). To indicate that 
$\mathbf{c}'$ dominates $\mathbf{c}''$, we use the notation
$\mathbf{c}' \succ \mathbf{c}''$.

\begin{definition}[Pareto Set]
\label{pers02.def:Pareto-set}
The {\em Pareto-optimal set} of $S$ for the application $b$ is the
set:
\[ \mathcal{P}(S,b) = \{ \mathbf{c} \in \mathcal{C}^*(S) \ | \nexists \ \mathbf{c}' \in \mathcal{C}^*(S), \mathbf{c}' \succ \mathbf{c} \} \]
\end{definition}
that is, the set of configurations $\mathbf{c} \in
\mathcal{C}^*(S)$ not dominated by any other \comment{feasible}
 configuration.
The configurations belonging to the Pareto-optimal set are called \emph{Pareto-optimal configurations}, while the {\em Pareto-optimal front} is their image, i.e. the set:
\[ \mathcal{P}_{F}(S,b) = \{ \mathbf{o} | \mathbf{o} = E(\mathbf{c},b), \mathbf{c} \in \mathcal{P}(S,b) \} \]

The aim of a Design Space Exploration (DSE) strategy is to give a
good approximation of the Pareto-optimal front for a system $S$ and an
application $b$, simulating as few configurations as possible.

%----------------------------------------------------------------

\comment{Mi sembra solo un cambio di notazione. }
On the contrary, we call ``parameter space representation'' of the
Pareto set the representation in which each configuration $\mathbf{p}$
is simply represented as a point of $P_{1}\times\dots\times P_{n}$
space. We will investigate the relation between the ``objective space
representation'' and the ``parameter space representation''. We
are particularly interested in the relation between the evolution
of the objective space representation of the Pareto set and the evolution
of its parameter space representation.

\begin{figure}[h]
\includegraphics[width=0.5\columnwidth]{img/Pareto_set_and_front}
\caption{Relation between Pareto set (its ``parameter space representation'')
and Pareto front (``objective space representation''). \comment{Questa figura sembra molto ``didattica''. La relazione tra Pareto front e Pareto set magari deve essere illustarta a studenti ma in un paper va bene? Io la apprezzerei se la trovassi, ma di solito si fa cosi'?}}
\end{figure}

\subsection{Sketch of the Idea}
\secL{sketch}

In the design process, there are often a lot of ``local decisions''
in which the designer can take advantage of well-known results or
heuristics. For example, if we consider a VLIW architecture in a configuration with very few registers, it is well-known that increasing the number of ALUs does not bring tangible better performances
but only cause an increase in area occupation. An intelligent exploration
algorithm would not waste time evaluating configurations with many
ALUs and few registers. If $p_{1}$ is the parameter ``number of
registers'' and $p_{2}$ is the parameter ``number of ALUs'', we
can say that the region:
\[
R=\left\{ \left.\left(v_{1},v_{2}\right)\in P_{1}\times P_{2}\right|v_{1}<s_{1},v_{2}>s_{2}\right\} 
\]
 (where $s_{1}$ and $s_{2}$ are some threshold values) is \emph{uninteresting}. This clearly shows that there are cases when, if a parameter lies in
a certain range, there are ranges of the other that are not worth exploring. Intuitively, the Cartesian product of these ranges is an ``uninteresting
 region''.

The problem is that, in complex scenarios, not all relations of this
kind are known in advance. Some dependency may exist but may be hidden
to the designer. Or they may be ``local'', i.e. two
parameters may be related by a dependency but only if their values
lie on some ranges (so we cannot treat them as they were related by
a ``complete'' dependency). 
Therefore, as stated in section \ref{sec:Related-work},
setting up an exploration trying to take into account as much dependencies
as possible is a hard task that is destined to be incompletely
accomplished. We need a methodology to ``automatically'' recognize
interesting or uninteresting regions without requiring a priori knowledge
to the designer, that is what our algorithm does.

We also consider, in measuring how much interesting a region is, the
\emph{innovation} that it adds. If, during a design space exploration,
a Pareto front is temporarily calculated, adding a new Pareto front
point near to the previous ones is not a considerable innovation,
because the way it fulfills the objectives is similar to the other
already evaluated configurations. On the contrary, adding a new Pareto
front point that is distant from previous ones is remarkable,
since it let the experimenter discover potential performance that have not yet considered before.

%TODO: sync name of values with formulation
\subsection{Formal definitions}
We start by defining the parameter interval and the concept of interval contiguity. Using these definition we construct the region and define the region contiguity. Then we define the operations of splitting and merging regions. Finally, we provide the definition of separation.

\begin{definition}[Parameter Interval] Let $p_{i}$ be a parameter and $a_{i},b_{i}\in V_{i},a_{i}<b_{i}$.
A $p_{i}$- interval is
\[
\left[a_{i}..b_{i}\right]=\left[a_{i},b_{i}\right]\cap P_{i}
\]
i.e. taking the interval $\left[a_{i}..b_{i}\right]$
is the set of values of $V_{i}$ lying in $\left[a_{i},b_{i}\right]$)
\end{definition}

We consider only parameters $p_{i}$ with ordering, i.e. such
that $\forall a,b\in V_{i}$ it is possible
to say $a<b$ or $a=b$ or $a>b$.

%TODO: DESCRIVERE COME SI PENSA DI TRATTARE GLI SCENARI IN CUI ESISTONO
%PARAMETRI SENZA ORDINAMENTO{]}

	\begin{figure}[t]
		\begin{center}
			\subfigure[Examples of contiguous and non contiguous intervals.]{
				\includegraphics[width=0.45\textwidth]{img/contiguous_intervals_or_not} }
			\subfigure[Examples of regions.]{
				\includegraphics[width=0.45\textwidth]{img/regions} }
		    \figLC{interval_and_region}{Intervals and regions.}
		\end{center}
	\end{figure}

\begin{definition}[Contiguous intervals]
\label{pers02.def:Contiguous-intervals}
Two intervals are contiguous iff they do not overlap and merging them results in a new interval.More formally, let $\left[a_{i}..b_{i}\right]$ and $\left[x_{i}..y_{i}\right]$
be two $p_{i}$-intervals. They are said to be contiguous (see \figR{interval_and_region}(a) ) iff
\begin{align}\begin{array}{l}
	\left[a_{i}..b_{i}\right]\cap\left[x_{i}..y_{i}\right]=\emptyset \mbox{      and} \\
	\left[a_{i}..b_{i}\right]\cup\left[x_{i}..y_{i}\right]$ is a $p_{i}$-
interval or $\left[x_{i}..y_{i}\right]\cup\left[a_{i}..b_{i}\right]
\end{array}\end{align}
\end{definition}

\begin{definition}[Region] 
A region is a cartesian product of parameter intervals, as illustrated in \figR{interval_and_region} (b). More formally, if $p_1,\dots,p_n$ are the parameters of the system, and $\left[ a_{i}..b_{i} \right]$ is a $p_i$-interval, a region is a set of the following form:
\[
R=\left[a_{1}..b_{1}\right]\times\dots\times\left[a_{n}..b_{n}\right]=\prod\left[a_{i}..b_{i}\right]
\]
\end{definition}

We will refer later to ``big'' and ``small'' regions. The sense
of these terms is related to the cardinality of regions.
\begin{definition}[Interval and region comparison]
A $p_{i}$-interval $\left[a_{i}..b_{i}\right]$
is said to be bigger then another $p_{i}$-interval $\left[c_{i}..d_{i}\right]$
iff 
$\left|\left[a_{i}..b_{i}\right]\right|>\left|\left[c_{i}..d_{i}\right]\right|$.
Similarly, a region $R_{1}$ is said to be \emph{bigger }then $R_{2}$, if
$\left|R_{1}\right|>\left|R_{2}\right|$
\end{definition}
Note that with $|\cdot|$ we indicate the cardinality of a set. Therefore, $\left|\left[a_{i}..b_{i}\right]\right|$ is the number of values of $V_i$ lying between $a_{i}$ and $b_{i}$ and has not to be confused with $b_i - a_i$.
Similarly $\left|R_{i}\right|$ is the number of configurations inside the region $R_i$.

We now provide some rules to split and merge regions. They will be
useful when explaining our proposed exploration algorithm.
\begin{definition}[Splitting an interval]
\label{pers02.def:Splitting-a-region}
Given a $p_i$ interval $\left[a_{i}..b_{i}\right]$ with more than one element, i.e. $\left|\left[a_{i}..b_{i}\right]\right|>1$, the splitting operation produces as result two contiguous intervals $\left[a_{i}..c_{i}\right],\left[d_{i}..b_{i}\right]$
 such that
	\begin{align}
		\begin{cases}
		\left[a_{i}..c_{i}\right]\cap\left[d_{i}..b_{i}\right] & =\emptyset\\
		\left[a_{i}..c_{i}\right]\cup\left[d_{i}..b_{i}\right] & =\left[a_{i}..b_{i}\right]\\
		\left|\left[a_{i}..c_{i}\right]\right| & =\left\lceil \frac{\left|\left[a_{i}..b_{i}\right]\right|}{2}\right\rceil 
		\end{cases}
	\end{align}

We define the split operator $\psi$ as the one that, applied to the intervals, gives the result of the split operation, i.e., in the case above, $\psi \left[a_{i}..b_{i}\right]=\lbrace \left[a_{i}..c_{i}\right],\left[d_{i}..b_{i}\right] \rbrace$.
If $\left|\left[a_{i}..b_{i}\right]\right|=1$, the interval
cannot be split and $\psi \left[a_{i}..b_{i}\right]=\left[a_{i}..b_{i}\right]$.
\end{definition}

The interval split is illustrated in \figR{split}~(a). Note that, when splitting an interval, the resulting intervals must satisfy two conditions: (i) they must be contiguous and, merging them, the initial interval must be obtained; (ii) the point in which we cut the interval is not arbitrary, but it is close to the middle.

\begin{definition}[Splitting a region]
\label{pers02.def:Splitting-a-region}
Consider a region $R=\prod\left[a_{i}..b_{i}\right]$.The split operator $\psi$ divides it in a set of subregions obtained splitting each $p_i$-interval and combining the resulting subintervals with the cartesian product, as in \figR{split}~(b). Formally:
	\begin{align}
		\psi(R) = \left\{ \left.\prod\left[x_{i}..y_{i}\right]\right|\left[x_{i}..y_{i}\right]\in \psi\left[a_{i}..b_{i}\right]  \right\} 
	\end{align}

\end{definition}

	\begin{figure}[t]
		\begin{center}
			\subfigure[Splitting an interval.]{
				\includegraphics[width=0.30\textwidth]{img/splitting_intervals} }
			\subfigure[Splitting a region.]{
				\includegraphics[width=0.60\textwidth]{img/splitting_regions} }
		    \figLC{split}{Split operations.}
		\end{center}
	\end{figure}

\begin{definition}[Contiguous regions]Let $R_{1}=\prod\left[a_{i}\dots b_{i}\right]$ and $R_{2}=\prod\left[x_{i}..y_{i}\right]$
be two regions. They are said to be contiguous iff they do not overlap and, merging them, a new region can be obtained. Formally, $R_1$ and $R_2$ are contiguous iff there exists a $j$ such that:
	\begin{align}\begin{array}{l}
		\left[a_{i}..b_{i}\right]=\left[x_{i}..y_{i}\right]$
	\mbox{ for } $i\neq j \mbox{      and} \\
		\left[a_{j}..b_{j}\right]$ and $\left[x_{j}..y_{j}\right] \mbox{ are contiguous intervals}
	\end{array}\end{align}
\noindent We call $p_j$ the \emph{contiguity parameter} for the two regions.
\end{definition}

Examples of contiguous and non contiguous regions are given in \figR{contiguous_regions}. 
\comment{Spiegare le figure e inserire a dx una figura con le contiguous regions mergiate. Le definizioni buttate cosi' rendono difficile la lettura a chi non sa ancora dove vogliamo andare a parare. Giustifica ogni definizione prima di darla. Ad es. quando parli di splittare, spiega che quando le regioni sono interessanti noi le splittiamo cosi' da esplorarle meglio. Al contrario quando le mergiamo. Mettere una sezione del tipo "Required operations at a glance" per giustificare tutte le definizioni prima di buttarle}.

	\begin{figure}[h]
		\includegraphics[width=0.5\columnwidth]{img/contiguous_regions}
		\figLC{contiguous_regions}{Contiguous and non contiguous
		regions.}
	\end{figure}

\begin{definition}[Merging intervals]
Let $\left[a_{i}\dots b_{i}\right]$ and $\left[x_{i}..y_{i}\right]$ be two contiguous $p_i$-intervals. The merge operation applied to them produces a new $p_i$-interval of the following form:
	\begin{align}
		\mu	 \left( \begin{array}{l}
				\left[a_{j}..y_{j}\right] \\
				\left[x_{j}..b_{j}\right]
		     \end{array} \right)
		=\begin{cases}
			\left[a_{j}..y_{j}\right] & \mbox{ if }b_{j}<x_{j}\\
			\left[x_{j}..b_{j}\right] & \mbox{ if }y_{j}<a_{j}
		\end{cases}
	\end{align}

\end{definition}

\begin{definition}[Merging regions]
\label{pers02.def:Merging-regions}
Let $R_{1}=\prod\left[a_{i}\dots b_{i}\right]$ and $R_{2}=\prod\left[x_{i}..y_{i}\right]$ be two contiguous regions and let $p_j$ be the contiguity parameter. 
The merge operation $\mu$ applied to $R_1$ and $R_2$ produces a new region of the following form:
	\begin{align}
		\mu(R_1,R_2)=\prod_{i<j}\left[a_{i}..b_{i}\right]
		\times
		\mu	 \left( \begin{array}{l}
				\left[a_{j}..y_{j}\right] \\
				\left[x_{j}..b_{j}\right]
		     \end{array} \right)
		\times
		\prod_{i>j}\left[a_{i}..b_{i}\right]
	\end{align}
\end{definition}

Note that $\mu(R_{1},R_{2})$ is mathematically equivalent to $R_{1}\cup R_{2}$. However, we use the notation above to enforce the requirement that
$R_{1}$ and $R_{2}$ must be to contiguous regions.

In \secR{sketch} we observed that, in order to estimate how much interesting is a region, it is crucial to understand the novelty that it introduces, i.e. how distant the Pareto-optimal configurations discovered in that region are from the Pareto-optimal configuration previously discovered. The concept of distance is expressed by the operator defined in the next definition. We prefer to call it \emph{separation} rather than \emph{distance}, because usually distance denote a commutative operator while our definition of separation is non commutative.

\begin{definition}[Separation]
Let $\mathbf{c}',\mathbf{c}''\in S$ be two configurations, $b\in\mathcal{B}$ a benchmark application and $\mathbf{o}',\mathbf{o}''\in \Re^m$ the representation of the configurations in the objective space, i.e. $\mathbf{o}'=E \left(\mathbf{c}', b\right)$, $\mathbf{o}''=E \left(\mathbf{c}'', b\right)$, where $E$ is the evaluation function. The separation between
$\mathbf{c}'$ and $\mathbf{c}''$ is 
	\[
	s\left(\mathbf{c}'\rightarrow\mathbf{c}''\right)=\sum_{i=1}^{m}\left|\frac{o'_{i}-o''_{i}}{o'_{i}}\right|
	\]
where $o'_i$ and $o''_i$ are the $i$-th components of $\mathbf{o}'$ and $\mathbf{o}''$, respectively.
\end{definition}

The separation between $\mathbf{c}'$ and $\mathbf{c}''$ measures how
much we must vary the image of the former to obtain the image of the latter. Note that the separation is a \emph{normalized} measure, thanks to $o'_i$ at the denominator. This is important since usually objectives are heterogeneous, e.g. delay and power consumption, having different unity of measurement and different scale. If we did not use a normalized measure, in presence of objectives whose absolute values are small and other objectives with big values, the notion of distance would have been biased toward the latter.