### 2014-07-28 paper.tex
Some fixes

%\documentclass[final,journal,letterpaper]{IEEEtran}
\documentclass[prodmode,acmtecs]{acmsmall}

\usepackage{subfigure}
\usepackage{cite}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{url}
\usepackage{balance}
\usepackage{float}
\usepackage{threeparttable}

\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{textcomp}
\usepackage{mathrsfs}
%\usepackage{amsthm}

\usepackage{mathptmx}
\usepackage[scaled=.90]{helvet}
\usepackage{courier}

\usepackage{color}

\usepackage{listings}
\usepackage[ruled]{algorithm2e}

\lstset{
   language=C,
   basicstyle=\small,
   keywordstyle=\bfseries,
   identifierstyle=\ttfamily,
   stringstyle=\ttfamily,
   numbers=left,
   numberstyle=\tiny,
   stepnumber=1,
   numbersep=-5pt,
   showstringspaces=false
%   frame=single %trbl%
}

\usepackage[normalem]{ulem}

%------------------------------------------------------------------------------
\newcommand{\figR}[1]{Fig.~\ref{fig:#1}}
\newcommand{\figLC}[2]{
        \caption{#2}
        \label{fig:#1}
        %\vspace{-10pt}
}
\newcommand{\secL}[1]{\label{sec:#1}}
\newcommand{\secR}[1]{Sec.~\ref{sec:#1}}

\newcommand{\defL}[1]{\label{def:#1}}
\newcommand{\defR}[1]{Def.~\ref{def:#1}}

\newcommand{\etal}{\emph{et al.}}
\newcommand{\eg}{\emph{e.g.}}
\newcommand{\ie}{\emph{i.e.}}
\newcommand{\etc}{\emph{etc.}}

\newcommand{\comment}[1]{\textcolor{red}{\bf #1}}
%\newcommand\comment[1]{} % When you enable this lines, all the notes will disappear

%\renewcommand{\baselinestretch}{0.98}\normalsize
%------------------------------------------------------------------------------
%\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
%\theoremstyle{plain}
%\newtheorem{thm}{\protect\theoremname}
%\theoremstyle{definition}
%\newtheorem{defn}{\protect\definitionname}
%\theoremstyle{remark}
%\newtheorem{rem}{\protect\remarkname}

%\usepackage{babel}
\providecommand{\definitionname}{Definition}
\providecommand{\remarkname}{Remark}
\providecommand{\theoremname}{Theorem}

\begin{document}
% Page heads
\markboth{V. Catania et al.}{Parameter Space Representation of Pareto Front to
Explore Hardware-Software Dependencies}

\title{Parameter Space Representation of Pareto Front to
Explore Hardware-Software Dependencies}
\author{VINCENZO CATANIA
\affil{University of Catania}
ANDREA ARALDO
\affil{University of Catania}
DAVIDE PATTI
\affil{University of Catania}
}

%\maketitle
\begin{abstract}
Embedded Systems Design requires conflicting objectives to be
optimized with an appropriate choice of hardware/software parameters.
A simulation campaign can guide the design in finding the best
trade-offs, but due to the big number of possible configurations it
is often unfeasible to simulate all of them. For these reasons Design Space
Exploration (DSE) algorithms aim at finding near-optimal system
configurations only simulating a subset of them.

In this work we present PS, a new multiobjective exploration
algorithm. The basic idea is to recognize \emph{interesting} regions, i.e.
regions of the configuration space that are proving to provide better
configurations with regard to other ones. PS evaluates more configurations in
interesting regions while exploring less thoroughly the rest of the
configuration space. After a detailed formal description of the
algorithm and the underlying concepts, we show a 
case study involving the hardware/software exploration of a VLIW
architcture. Qualitative and quantitative comparisons of PS
against a well-know multiobjective genetic approach demonstrate that,
while not outperforming it in terms of dominance, the proposed
approach can trade the uniformity and granularity qualities of the
solutions found for obtaining larger Pareto sets, thus representing a
further choice for designer when objective constrains allow it.
\end{abstract}

\category{D.2.2}{Design Tools and Techniques}{}

\terms{Design, Algorithms, Performance}

\keywords{DSE, MultiObjective Optimization, Genetic Algorithms}

\acmformat{Vincenzo Catania, Andrea Araldo and Davide Patti, 2014. Parameter Space Representation of Pareto Front to
Explore Hardware-Software Dependencies.}

\begin{bottomstuff}
Author's addresses: V. Catania, DIEEI University of Catania;
A. Araldo, DIEEI University of Catania;
D. Patti, DIEEI University of Catania.
\end{bottomstuff}

\maketitle
%------------------------------------------------------------------------------
\input{introduction.tex}
\input{concepts.tex}
\input{algorithm.tex}
\input{experiments.tex}
%------------------------------------------------------------------------------

\section{Conclusions}
In this work we presented a multiobjective strategy which 
introduces the concept of Parameter Space representation of Pareto
Front. A case study of a VLIW architecture involving strong
hardware/software dependencies has been carried out to evaluate
effectiveness with different amount of simulations budget.
A qualitative/quantitative comparison against a widespread multiobjective genetic
approach showed how the proposed strategy can result in a different
distribution of Pareto points which trades solutions granularity with
improved pareto extension.

\balance

\bibliographystyle{ACM-Reference-Format-Journals} 
\bibliography{bibliography}

% TODO: G2012 bib not used ?

\end{document}