### 2014-10-27 experiments.tex
Draft of Comparison Setup subsection

\section{A Case Study: VLIW Hardware/Software Design Space}
\secL{ee}
Although PS has been formally described as a general exploration algorithm, the goal of this work is giving a method applicable to the embedded system design. For this reason, we evaluate PS in this context. To this aim, we use a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item \emph{Multi-Objective trade-offs}: VLIW architectures allow designers
to trade between power, energy and performances objectives, resulting
in extended Pareto sets which are an ideal ground for our testing
purposes focused on parameter-space representation
\item \emph{Compiler awareness}: code scheduling for the execution of
applications is statically obtained by the compiler which
performs code transformations ``being aware'' of the underlying hardware
configuration.  This tight hardware/software dependence makes the
VLIW scenario perfectly suitable for investigating hardly predictable
parameter interactions.
\end{itemize}

In this section, we will also briefly describe the parameterized platform used
as testbed for the experiments, the general evaluation flow along with
the high-level estimation models and applications used as benchmarks.

\subsection{Design Space: Hardware and Software Parameters}
Conceptually, system parameters can be classified in three main categories:
\emph{processor}, \emph{memory sub-system} and \emph{compiler}. The
first two categories are directly related to the physical
implementation of the system, so parameters of those
classes are usually referred ``hardware parameters''. On the other
side, all the parameters affecting the behaviour of the compiler
are referred as ``software parameters'', being directly involved in
the process of generating the application executables running on the
underlying hardware.

Processor hardware parameters are directly related to the VLIW core and includes the size
of the register files and the number of functional
units of each type. In particular, five different types of register
files are supported:
GPR (32-bit registers for integers), FPR (64-bit registers for
floating point values) PR (1-bit registers used to store the Boolean
values of predicated instructions), BTR (64-bit registers containing
information about possible future branches) and CR (32-bit control
registers containing information about the internal state of the
processor). The functional units involved are: \emph{Integer units},
\emph{floating point units}, \emph{memory units} (associated with
load/store operations) and \emph{branch units} (associated with branch
operations). With respect to the memory sub-system hardware, the parameters
include \emph{size}, \emph{associativity} and
\emph{block size} for each of the three caches: First-level data cache
(L1D), first-level instruction cache (L1I) and second-level unified
cache (L2U).

On the software side, a set of the most impacting code
compilation parameters has been selected and included in the
configuration space, in particular:
\begin{itemize}
\item \emph{tcc\_region}: Specifies the scope of action of the compiler and the type
of code transformation involved (basic block, super block and hyper
block) 
\item \emph {max\_unroll\_allowed}: The number of unroll iterations allowed
\item \emph{regroup\_only}: Avoids inlining 
\item \emph{do\_classic\_opti}: A set of classical optimization, not VLIW related,
such as common expression removal and loop unrolling
\item \emph{do\_prepass\_scalar\_scheduling}: Performs a schedule before
forming regions 
\item \emph{do\_postpass\_scalar\_scheduling}: Performs a schedule after the region formation 
\item \emph{do\_modulo\_scheduling}: Modular scheduling of loop code
\item \emph{memvr\_profiled}: Performs a memory-dependencies profiling 
\end{itemize}

As regards the class of benchmark being considered, a set applications representative of some common
frequently running code kernels has been selected. Table~\ref{tab:bench} shows the set of applications
along with a brief description.
% IEEE table style
%\begin{table}
%	\centering
%	\caption{Benchmarks}
%	\label{tab:bench}
%	\begin{tabular}{ll}
%	\hline
%	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
%	\hline
%	Alloca\_test & Memory array allocation test \\
%	Bmm & Matrices multiplication and elements sum \\
%	Fir\_int & Finite impulse response \\
%	Mm & Floating point matrices multiplication \\
%	Sha256 & Cryptocurrency header Hashing \\
%	Wave & Wavefront computation \\
%	\hline
%	\end{tabular}
%\end{table}

% ACM table style
\begin{table}
	\tbl{Benchmarks\label{tab:bench}}{
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}}
\end{table}

%------------------------------------------------------------------------------
\subsection{Configuration Evaluation Flow}

To evaluate and compare the performance indexes of different
architectures for a specific application, one needs to simulate the
architecture running the code of the application. To make
architectural exploration possible both the compiler and the simulator
have to be retargetable. The open source project Trimaran~\cite{trimaran_hp} provides these
tools and together with estimation framework \ee~\cite{palpatti_estimedia03} has been
adopted as a tested and flexible platform for carrying out the
experiments described in the previous subsection.

We will now show a functional scheme highlighting the main blocks of
the \ee{} framework and the interface with the Trimaran tools. The
input for the whole evaluation flow is the source of the application
involved in the optimization and the configuration of the architecture
being evaluated.  With reference to Figure~\ref{fig:evaluation_flow}
this input is represented by the blocks \emph{App.c} and
\emph{Config}.
\begin{figure}
        \centering
        \includegraphics[width=0.65\textwidth]{pictures/evaluation_flow.eps}
        \caption{Block diagram of the framework.}
        \label{fig:evaluation_flow}
\end{figure}

The application (App.c) is first compiled by the Trimaran's compiler
front-end (IMPACT). At the end of the compilation flow,
Trimaran supplies a simulation library (Emulib) which makes it
possible to execute the VLIW code produced by Elcor, generating a file
(Stats) containing the execution statistics (e.g., instruction mix,
execution cycles, utilization of functional units, etc.). A cache
simulator, along with a bus simulator, is used to gather information
about the behavior of the memory hierarchy in terms of miss rate and
data/address traffic on the interconnection buses.

Together with the configuration of the system, the statistics produced
by simulation contain all the information needed to apply the area,
performance and power consumption estimation model implemented in the
\emph{Estimator} component of \ee{}. The results obtained by these
models are the input for the \emph{Explorer} component. This component
executes an optimization algorithm, the aim of which is to modify the
parameters of the configuration so as to minimize the three cost
functions (area, execution time and energy/power consumption).

\subsection{Estimation Models}
Several models have been implemented in the \ee{} platform in order to
estimate the power/energy/performance indexes of each simulated system
configuration.

The average power consumed by the processor was estimated using an
adaptation of the Cai-Lim model~\cite{cai_micro99} to the VLIW
processor. As regards the cache subsystem, a transition-based model
was used, according to the equations described
in~\cite{kamble_islped97}.  The main memory energy is based on the
model in~\cite{shiue_dac99} and assumes a per main memory access
energy of $4.95 \times 10^{-9} J$ based on the data for the Cypress
CY7C1326-133 memory chip.  The contribution towards power consumption
made by the interconnection system was calculated by counting the
number of transitions on the bus lines and applying the formula
$P_{bus} = 1/2 V_{dd}^2 \alpha f C_l$ where $V_{dd}$ is the supply
voltage, $\alpha$ is the switching activity, $f$ is the clock
frequency and $C_l$ is the capacity of a bus line. For the on-chip
buses we also considered the coupling capacitances between bus lines,
using the model in~\cite{lekatsas_dac01}. For a detailed description
of the models used and their adaption to the case of a VLIW based
system see~\cite{palpatti_estimedia03}.

\subsection{Comparison Setup}

The results presented in the next section compare PS with one widely
adopted Multi-objective Genetic Algorithm~\cite{knowles_techrep06}. To
the best of our knowledge, genetic approaches in general still
represent the best compromise between efficiency (i.e. time required
for exploration) and accuracy of the reported Pareto Sets.  Note that,
for the reasons described above, mono-objective approaches have been
explicitly excluded from this analysis.  We also exclude other
approaches (e.g.  dependency based~\cite{givargis_tvlsi02}) that make
use of some a-priori knowledge about the role and meaning of each
parameter, since our aim is evaluate the capacity of focusing on
interesting parameter regions without any external judgement or help
based on heuristics.  In particular in this work when referring to
genetic approach (or GA) we will denote the widely spread 
\emph{SPEA}~\cite{zitzler_eurogen01} variant of the Multi-objective
Genetic Algorithm, remanding to
works like~\cite{zitzler_ec00}~\cite{zitzler_tec03} for a detailed
analysis on how it could differ from other genetic approaches and what
could be expected from similar multi-objetive strategies.

\comment{TODO controllare inglese}
Even if an analogy between the concepts of ``era'' and ``generation'' 
can be considered, a fixed amount of total different simulations has been
considered for each of the tests. The purpose is to make the
comparison fair from a algorithm performance perspective, since a
fixed budget of simulations is directly related to the amount of time
required to execute the exploration; on the other side, simply
referring to number of generations/era could be misleading when a
significative number of configuration remain unaltered in the current
population (e.g. because of a local optimum).

The set of default algorithm tuning parameters used for both the
genetic and PS approaches to perform the simulations are shown in
Table~\ref{tab:defaults}. Given the similarity, both in terms of
meaning and functionality, of the parameter $K$ with the population size
of the genetic approach, they both have been set to the same values. As discussed
when introducing the main PS algorithm concepts, an accurate
investigation of algorithm tuning is beyond the scope of this
introductory work. Nevertheless, a set of preliminary tests with
different values of $K$ have been conducted. As shown in
Figure~\ref{fig:kimpact}, which refer to the alloca\_test application
and a fixed amount of 500 configurations, initially increasing values
of $K$ seem to improve both pareto (left picture) and objective
distribution (right picture), but after some threshold the $K$ does
not seem to provide any benefit, even returning to slightly worse
results for the last value considered (100). Similar results have been
reported on average for the different benchmarks considered, which are
not included for space reasons. A complete resource with the whole 
results and data obtained in the tests performed can be found
at~\cite{ps_results}.

\begin{table}
	\tbl{Benchmarks\label{tab:defaults}}{
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Default Value} \\
	\hline
	GA population size & 30 \\
	GA crossover probability & 0.8 \\
	GA mutation probability  & 0.1 \\
	$K$ era budget & 30 \\
	alpha threshold & 1 \\
	Total simulations budgets & 100-500 \\
	\hline
	\end{tabular}}
\end{table}

\begin{figure}
  \figLC{kimpact}{Impact of increased K values on Pareto fronts and distribution for the alloca\_test application.}
  \begin{center}
    \includegraphics[width=0.45\textwidth]{pictures/alloca_K.eps} 
    \includegraphics[width=0.45\textwidth]{pictures/alloca_K_ALL.eps} 
  \end{center}
\end{figure}

%------------------------------------------------------------------------------

\subsection{Evaluation metrics}
In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. Both of them permit to quantify the extent of the Pareto front that is an important factor to consider when evaluating the quality of a Pareto front. Indeed, authors in \cite{zitzler_ec00,weise2012evolutionary} observe that an exploration algorithm should preferably provide Pareto fronts whose objective function range is large. In other words, following the terminology of~\cite{weise2012evolutionary}, a \emph{uniform convergence} should be guaranteed.
Pareto fronts with large objective function ranges permit to investigate in a broader way the potentials of the designed device. This is important, in general, since if only a narrow portion of the Pareto front is provided due to non-uniform convergence, the decision maker may neglect some important options~\cite{weise2012evolutionary}. Having a wide Pareto front becomes essential, in particular, when designing a device that may be required to satisfy different constraints. For example, the decision maker may want to design two different multi-purpose processors. The first design may adapt to situations in which the processor is used for real time elaboration (thus requiring low execution time) and is attached to the powerline (thus not having stringent power dissipation constraints). The second design, on the contrary, may fit situations in which the energy is provided by a battery (stringent power constraints) but only background computations are performed (long execution time is tolerable). These two processors may be successfully designed maintaining the same architecture, using different values for the design parameters. A wide Pareto front helps in achieving this goal.

For a given objective, the \emph{variation range} represents the ratio between the maximum and the minimum value observed for that objective.
\comment{AA: Per rappresentare l'extent del Pereto front, forse sarebbe meglio evitare di usare il variation range, per come l'hai definito tu, e usare la funzione M3 di \cite{zitzler_ec00}, Sec.5 ``Metrics of performance'', che mi pare rappresenti la stessa cosa, concettualmente. Questo ci permette di essere meno attaccabili (del tipo ``non te la prendere con noi, prenditela con Zitzler (che e' citato 3064 volte!) se non capisci l'utilita' della metrica'').}

The average \emph{normalised absolute dispersion error} measures
the average absolute difference between the distribution of points in
the objective space and an ideal distribution in which the points are
uniformly distributed over the objective space. 
Formally, let $O$ be
the image, in the objective space, of the configurations visited by
the design space exploration. The generic element of $O$ (\ie, a
solution) is a pair $(p,t)$ where $p$ and $t$ are the average power
and execution time, respectively. The two-dimensional objective space
is then partitioned by a $M_x \times M_y$ mesh. For each tile $T_i,
\ i=1, 2, \ldots, M_xM_y$ of the mesh, let $N_i$ be the number
of points in $O$ which fall in $T_i$. The average absolute error, $E_i$, for
$T_i$ is the absolute value of the difference between $N_i$ and the
ideal number of solutions, $\overline{N}$, which should fall in $T_i$
in case of uniform distribution. Such $\overline{N}$ can be simply
computed as the ratio between the cardinality of $O$ and the number of
tiles. Thus,
\[ E_i = |N_i - \overline{N}|, \]
where $\overline{N} = |O| / (M_x M_y)$. The average
normalised absolute dispersion error ($ANADE$) is the average of $E_i$
normalised to the maximum absolute error $E_{max}$:
\[ ANADE = \frac{\sum_{i=1}^{M_xM_y} E_i/(M_xM_y)}{E_{max}}, \]
where $E_{max}$ can be computed as the average absolute error in the worst
case in which all the solutions fall in a single tile:
\[ E_{max} = \frac{(M_x M_y - 1) \overline{N} + |\overline{N} -
    |O|| }{M_x M_y}. \]

\subsection{Impact of the algorithm parameters}
\secL{parameter_impact}
\comment{IMPACT OF $K$. Si potrebbero presentare un plot in cui il numero totale di simulazioni e' fissato e vengono eseguiti diversi run dell'algoritmo con diversi valori di K. Per ogni run, il pareto set e' rappresentato nel plot, sottoforma di una curva diversa. Mi aspetterei che tutte le curve siano molto simili. Per ciascun run, il valore di variation range e average absolute dispersion error sono riportati. Mi aspetto valori simili, il che confermerebbe che la diversity nel nostro algoritmo non dipende da $K$. Ne vale la pena rifare questi run? Ci vuole troppo tempo?}

\comment{Si potrebbe mettere un plot con nelle ascisse l'era e nelle ordinate il numero di split avvenuti in quell'era. Si potrebbero disegnare diverse curve, una per ogni valore di alpha, per far vedere che il numero di split e' piu' alto q. Recommended values are close to 1, as results in \secR{ee} confirm. Using different values, if sufficiently close to 1, does not have a considerable impact on the resulting Pareto front approximation. Ne vale la pena rifare questi run? Ci vuole troppo tempo?}

\subsection{Results}

\figR{pareto_fronts_100} and \figR{pareto_fronts_500} show the resulting Pareto fronts for two
different scenarios of 100 and 500 simulations budget. \comment{AA: per curiosita': quando facevi piu' run, PS andava ancora peggio rispetto a GA?} When possible,
similar algorithm parameters have been adopted. In particular a
population size of 30 has been chosen for the genetic approach (GA), corresponding to the
same value for the $K$ of PS. \comment{AA: che valore di $alpha$ hai usato?} The number of generations (or era)
has been chosen so that the only the given amount of simulations is
actually performed. Thus, when accounting these simulations only unique simulations
has been considered, removing repetitions and unfeasible
configurations.

% IEEE table style
%\begin{table}
%  \centering
%  \begin{tabular}{ccc}
%    \includegraphics[width=0.30\textwidth]{pictures/alloca_100.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/bmm_100.eps} & 
%    \includegraphics[width=0.30\textwidth]{pictures/fir_int100.eps} \\
%    \includegraphics[width=0.30\textwidth]{pictures/mm_100.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/sha_100.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/wave_100.eps} 
%  \end{tabular}
%  \caption{Pareto fronts found by PS and GA for a fixed budget of 100 configurations.}
%  \label{fig:pareto_fronts_100}
%\end{table}
%
%\begin{table}
%  \centering
%  \begin{tabular}{ccc}
%    \includegraphics[width=0.30\textwidth]{pictures/alloca_500.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/bmm_500.eps} & 
%    \includegraphics[width=0.30\textwidth]{pictures/fir_int500.eps} \\
%    \includegraphics[width=0.30\textwidth]{pictures/mm_500.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/sha_500.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/wave_500.eps} 
%  \end{tabular}
%  \caption{Pareto fronts found by PS and GA for a fixed budget of 500 configurations.}
%  \label{fig:pareto_fronts_500}
%\end{table}

\begin{figure}
  \figLC{pareto_fronts_100}{Pareto fronts found by PS and GA for a
  fixed budget of 100 configurations. \comment{AA: nelle due figure
  centrali la linea di GA e' continua, mentre in tutte le altre figure
  e' tratteggiata. Il revisore ci chiede di uniformare. Non so come
  riottenere le figure. Potresti pensarci tu?}}
  \begin{center}
    \includegraphics[width=0.30\textwidth]{pictures/alloca_100.eps}
    \includegraphics[width=0.30\textwidth]{pictures/bmm_100.eps}
    \includegraphics[width=0.30\textwidth]{pictures/fir_int100.eps}
    \includegraphics[width=0.30\textwidth]{pictures/mm_100.eps}
    \includegraphics[width=0.30\textwidth]{pictures/sha_100.eps}
    \includegraphics[width=0.30\textwidth]{pictures/wave_100.eps} 
  \end{center}
\end{figure}

\begin{figure}
  \figLC{pareto_fronts_500}{Pareto fronts found by PS and GA for a
  fixed budget of 500 configurations.\comment{AA: nelle due figure
  centrali la linea di GA e' continua, mentre in tutte le altre figure
  e' tratteggiata. Il revisore ci chiede di uniformare. Non so come
  riottenere le figure. Potresti pensarci tu?}}
  \begin{center}
    \includegraphics[width=0.30\textwidth]{pictures/alloca_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/bmm_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/fir_int500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/mm_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/sha_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/wave_500.eps} 
  \end{center}
\end{figure}

Low budget Pareto sets are shown in \figR{pareto_fronts_100}. With a so small number of
configuration being investigated the two sets are often overlapping,
being also strongly dependent upon the application being considered.
These results are obtained in about 5 eras and demonstrate that with a so limited budget PS can compete with
genetic approach while even discovering very different Pareto fronts. 
Considering the extended simulation budget as depicted in
\figR{pareto_fronts_500}, we can qualitatively observe that
even if PS does not strictly outperforms genetic in terms of dominance, it shows in
some cases largely Pareto in terms of extension. We can intuitively
relate this to the ``novelty-based'' score system, so that instead of
focusing on getting better points inside a given range it tries to
enlarge the range itself, resulting in less dense but more extended
Paretos. Of course we cannot classify this different behaviour as
``better'' or ``worse'' as compared to the genetic one, it depends on
various design factors such as desired results granularity, prefixed objective
range constraints, estimated error of the models used etc.

A comparison between the variation ranges of both power dissipation
and execution time, obtained by PS and GA for different benchmarks is
shown in \figR{dispersion}(a) and (b) respectively.  As it can
be observed, the PS exploration provides solutions which fall on a
range that is, on average, wider than the one provided by a GA
exploration for power dissipation and execution time, respectively.
%\begin{table}
%  \centering
%  \begin{tabular}{c}
%    \includegraphics[width=0.7\textwidth]{pictures/variation_power.eps} \\
%    (a) \\
%    \includegraphics[width=0.7\textwidth]{pictures/variation_etime.eps} \\
%    (b) \\
%    \includegraphics[width=0.7\textwidth]{pictures/dispersion.eps} \\
%    (c) 
%  \end{tabular}
%  \caption{(a)(b) Variation range obtained by PS and GA for different benchmarks
%  and average dispersion absolute dispersion (c)}
%  \label{fig:dispersion}
%\end{table}

\begin{figure}
  \figLC{dispersion}{(a)(b) Variation range obtained by PS and GA for different benchmarks and average dispersion absolute dispersion (c). \comment{AA: manca il titolo nella figura (c)} }
  \begin{center}
    \subfigure[]{\includegraphics[width=0.7\textwidth]{pictures/variation_power.eps} }
    \subfigure[]{\includegraphics[width=0.7\textwidth]{pictures/variation_etime.eps} }
    \subfigure[]{\includegraphics[width=0.7\textwidth]{pictures/dispersion.eps} }
  \end{center}
\end{figure}

\figR{dispersion}(c) shows the average normalised absolute
dispersion errors for different benchmarks for PS and GA
exploration. As it can be observed, GA shows a tendency in
discovering more uniformly dispersed points (i.e. higher values in the
chart). These results,
together with the variation ranges of \figR{dispersion}(a) and
(b) confirm the of density vs extension trade-offs
when comparing PS and GA approaches. Again, it
is worth pointing out how a definitive claim about the superiority of
the PS behaviour is strictly dependent on the design 
requirements and constraints, e.g. whether we need wide or dense
results. Nevertheless, the proposed algorithm introduces a new
perspective on the allocation of a simulations budget, paving the way for
the development of further strategies for addressing the problem of
multi-objective hardware/software design space exploration.