### 2014-10-29 experiments.tex
Added table on design space

\section{A Case Study: VLIW Hardware/Software Design Space}
\secL{ee}
Although PS has been formally introduced and described as a general
multi-objective optimization algorithm, in this work we will
explicitly refer to the more restricted (but still very large)
embedded system design scenario,  since its properties
are the one that suggested to the authors the concepts of pareto-based
innovation and region scoring system.

In particular, we will use a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item \emph{Multi-Objective trade-offs}: VLIW architectures allow designers
to trade between power, energy and performances objectives, resulting
in extended Pareto sets which are an ideal ground for our testing
purposes focused on parameter-space representation
\item \emph{Software Compiler awareness}: code scheduling for the execution of
applications is statically obtained by the compiler which
performs code transformations ``being aware'' of the underlying hardware
configuration.  This tight hardware/software dependence makes the
VLIW scenario perfectly suitable for investigating hardly predictable
parameter interactions.
\end{itemize}

In this section, we will also briefly describe the parameterized platform used
as testbed for the experiments, the general evaluation flow along with
the high-level estimation models and applications used as benchmarks.

\subsection{Design Space: Hardware and Software Parameters}
\secL{designspace}
Conceptually, system parameters can be classified in three main categories:
\emph{processor}, \emph{memory sub-system} and \emph{compiler}. The
first two categories are directly related to the physical
implementation of the system, so parameters of those
classes are usually referred as ``hardware parameters''. On the other
side, all the parameters affecting the behaviour of the compiler
are referred as ``software parameters'', being directly involved in
the process of generating the application executable running on the
underlying hardware.

Processor hardware parameters are directly related to the VLIW core
and includes the size of the register files and the number of
functional units of each type. On the software side, a set of the most
impacting code compilation parameters has been selected.
Table~\ref{tab:param} show the complete set of parameters included in
the design space and their possible values. The total size of design
space, which grows with the product of parameter cardinalities is
about $1.33 \times 10^{14}$.

As regards the class of benchmark being considered, a set applications representative of some common
frequently running code kernels has been selected. Table~\ref{tab:bench} shows the set of applications
along with a brief description.
% IEEE table style
%\begin{table}
%	\centering
%	\caption{Benchmarks}
%	\label{tab:bench}
%	\begin{tabular}{ll}
%	\hline
%	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
%	\hline
%	Alloca\_test & Memory array allocation test \\
%	Bmm & Matrices multiplication and elements sum \\
%	Fir\_int & Finite impulse response \\
%	Mm & Floating point matrices multiplication \\
%	Sha256 & Cryptocurrency header Hashing \\
%	Wave & Wavefront computation \\
%	\hline
%	\end{tabular}
%\end{table}

\begin{table}
	\tbl{Design Space\label{tab:param}}{
	\begin{tabular}{lll}
	\hline
	\multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Values} & \multicolumn{1}{c}{Type}\\
	\hline
	General Purpose Registers & 16, 32, 64, 128 & HW \\
	Floating Point Registers & 16, 32, 64, 128 & HW \\
	Predicate Instructions registers & 32, 64, 128 & HW \\
	Branch Target Registers & 16, 32, 64 & HW \\
	Control Registers & 32, 64, 128 & HW \\
	\hline
	Integer Functional Units & 1, 2, 3, 4, 5, 6 & HW \\
	Floating Point Units & 1, 2, 3, 4, 5, 6 & HW\\
	Memory Units & 1, 2, 3, 4  & HW \\
	Branch Units & 1, 2, 3, 4  & HW \\
	\hline
	L1 Data Cache Size (KB) & 1, 2, 4, 8, 16, 32, 64, 128 & HW \\
	L1 Data Cache Blocksize & 32, 68, 128 & HW \\
	L1 Data Cache Associativity (KB) & 1, 2, 4 & HW \\
	L1 Instruction Cache Size (KB) & 1, 2, 4, 8, 16, 32, 64, 128 & HW \\
	L1 Instruction Cache Blocksize & 32, 68, 128 & HW \\
	L1 Instruction Cache Associativity (KB) & 1, 2, 4 & HW \\
	L2 Unified Cache Size (KB) & 32, 64, 128, 256, 512  & HW\\
	L2 Unified Cache Blocksize & 32, 68, 128 & HW \\
	L2 Unified Cache Associativity (KB) & 2, 4, 8 & HW \\
	\hline
        tcc\_region: Scope of action of the compiler & basic block, super block, hyper block & SW \\
        max\_unroll\_allowed: unroll iterations allowed & none, medium, high & SW \\
        regroup\_only: Avoids inlining & yes, no & SW \\
        do\_classic\_opti: Classical optimizations & yes, no & SW \\
        do\_prepass\_scalar\_scheduling: Performs a prepass schedule & yes, no & SW \\
        do\_postpass\_scalar\_scheduling: Performs a schedule after & yes, no & SW \\
	do\_modulo\_scheduling: Modular scheduling of loop code & yes, no & SW \\
        memvr\_profiled: Memory-dependencies profiling & yes, no & SW \\
	\hline
	\end{tabular}}
\end{table}

% ACM table style
\begin{table}
	\tbl{Benchmarks\label{tab:bench}}{
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Benchmark} & \multicolumn{1}{c}{Application} \\
	\hline
	Alloca\_test & Memory array allocation test \\
	Bmm & Matrices multiplication and elements sum \\
	Fir\_int & Finite impulse response \\
	Mm & Floating point matrices multiplication \\
	Sha256 & Cryptocurrency header Hashing \\
	Wave & Wavefront computation \\
	\hline
	\end{tabular}}
\end{table}
%------------------------------------------------------------------------------
\subsection{Configuration Evaluation Flow}

To evaluate and compare the performance indexes of different
architectures for a specific application, one needs to simulate the
architecture running the code of the application. To make
architectural exploration possible both the compiler and the simulator
have to be retargetable. The open source project Trimaran~\cite{trimaran_hp} provides these
tools and together with estimation framework \ee~\cite{palpatti_estimedia03} has been
adopted as a tested and flexible platform for carrying out the
experiments described in the previous subsection.

We will now show a functional scheme highlighting the main blocks of
the \ee{} framework and the interface with the Trimaran tools. The
input for the whole evaluation flow is the source of the application
involved in the optimization and the configuration of the architecture
being evaluated.  With reference to Figure~\ref{fig:evaluation_flow}
this input is represented by the blocks \emph{App.c} and
\emph{Config}.
\begin{figure}
        \centering
        \includegraphics[width=0.65\textwidth]{pictures/evaluation_flow.eps}
        \caption{Block diagram of the framework.}
        \label{fig:evaluation_flow}
\end{figure}

The application (App.c) is first compiled by the Trimaran's compiler
front-end (IMPACT). At the end of the compilation flow,
Trimaran supplies a simulation library (Emulib) which makes it
possible to execute the VLIW code produced by Elcor, generating a file
(Stats) containing the execution statistics (e.g., instruction mix,
execution cycles, utilization of functional units, etc.). A cache
simulator, along with a bus simulator, is used to gather information
about the behavior of the memory hierarchy in terms of miss rate and
data/address traffic on the interconnection buses.

Together with the configuration of the system, the statistics produced
by simulation contain all the information needed to apply the area,
performance and power consumption estimation model implemented in the
\emph{Estimator} component. These are high-level abstraction models
that allow to quickly get an estimation while still providing a
discrete level of accuracy (for a detailed description of the models
see also~\cite{palpatti_estimedia03}).

Finally, the results obtained by these
models are the input for the \emph{Explorer} component. This component
executes an optimization algorithm, the aim of which is to modify the
parameters of the configuration so as to minimize the three cost
functions (area, execution time and energy/power consumption).

%\subsection{Estimation Models}
%
%The average power consumed by the processor was estimated using an
%adaptation of the Cai-Lim model~\cite{cai_micro99} to the VLIW
%processor. As regards the cache subsystem, a transition-based model
%was used, according to the equations described
%in~\cite{kamble_islped97}.  The main memory energy is based on the
%model in~\cite{shiue_dac99} and assumes a per main memory access
%energy of $4.95 \times 10^{-9} J$ based on the data for the Cypress
%CY7C1326-133 memory chip.  The contribution towards power consumption
%made by the interconnection system was calculated by counting the
%number of transitions on the bus lines and applying the formula
%$P_{bus} = 1/2 V_{dd}^2 \alpha f C_l$ where $V_{dd}$ is the supply
%voltage, $\alpha$ is the switching activity, $f$ is the clock
%frequency and $C_l$ is the capacity of a bus line. For the on-chip
%buses we also considered the coupling capacitances between bus lines,
%using the model in~\cite{lekatsas_dac01}.

\subsection{Comparison Setup}
\secL{comparison}

The results presented in the next section compare PS with one widely
adopted Multi-objective Genetic Algorithm~\cite{knowles_techrep06}. To
the best of our knowledge, genetic approaches in general still
represent the best compromise between efficiency (i.e. time required
for exploration) and accuracy of the reported Pareto Sets.  Note that,
for the reasons described above, mono-objective approaches have been
explicitly excluded from this analysis.  We also exclude other
approaches (e.g.  dependency based~\cite{givargis_tvlsi02}) that make
use of some a-priori knowledge about the role and meaning of each
parameter, since our aim is evaluate the capacity of focusing on
interesting parameter regions without any external judgement or help
based on heuristics.  In particular in this work when referring to
genetic approach (or GA) we will denote the widely spread 
\emph{SPEA}~\cite{zitzler_eurogen01} variant of the Multi-objective
Genetic Algorithm, remanding to
works like~\cite{zitzler_ec00}~\cite{zitzler_tec03} for a detailed
analysis on how it could differ from other genetic approaches and what
could be expected from similar multi-objetive strategies.

Even if an analogy between the concepts of ``era'' and ``generation'' 
can be considered, a fixed amount of total different simulations has been
considered for each of the tests. The purpose is to make the
comparison fair from a algorithm performance perspective, since a
fixed budget of simulations is directly related to the amount of time
required to execute the exploration; on the other side, just simply
referring to a given number of generations/era could be misleading when a
significative number of configuration remain unaltered in the current
population (e.g. because of a local optimum).

The set of default algorithm tuning parameters used for both the
genetic and PS approaches to perform the simulations are shown in
Table~\ref{tab:defaults}. Given the similarity, both in terms of
meaning and functionality, of the parameter $K$ with the population size
of the genetic approach, they both have been set to the same values. As discussed
when introducing the main PS algorithm concepts, an accurate
investigation of algorithm tuning is beyond the scope of this
introductory work. Nevertheless, a set of preliminary tests with
different values of $K$ have been conducted, similar the one shown in
Figure~\ref{fig:kimpact} which refer to the alloca\_test application
and a fixed amount of 500 configurations. We found that initially increasing values
of $K$ seem to improve both pareto (as in the picture on the left) and objective
distribution (right picture), but after some threshold value the $K$ does
not seem to provide any benefit, even returning to slightly worse
results for the last value of $K$ considered (100). Similar trends have been
found for the other benchmarks (not included for space reasons),
showing on average a value of $K$ around 30 as
reasonable sweet spot for the objectives. A complete online resource with the whole 
results and data obtained in the tests performed can be found
at~\cite{ps_results}.

\begin{table}
	\tbl{Benchmarks\label{tab:defaults}}{
	\begin{tabular}{ll}
	\hline
	\multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Default Value} \\
	\hline
	GA population size & 30 \\
	GA crossover probability & 0.8 \\
	GA mutation probability  & 0.1 \\
	$K$ era budget & 30 \\
	alpha threshold & 1 \\
	Total simulations budgets & 100-500 \\
	\hline
	\end{tabular}}
\end{table}

\begin{figure}
  \figLC{kimpact}{Impact of increased K values on Pareto fronts and distribution for the alloca\_test application.}
  \begin{center}
    \includegraphics[width=0.45\textwidth]{pictures/alloca_K.eps} 
    \includegraphics[width=0.45\textwidth]{pictures/alloca_K_ALL.eps} 
  \end{center}
\end{figure}

%------------------------------------------------------------------------------

\subsection{Evaluation metrics}
Qualitative evaluation of paretos is probably the simplest way for
having a quick glance on algorithm behviour and results. However, in
tests involving a comparison between different exploration strategies
that approach could not help in capturing the whole picture, e.g. missing
some important properties related to the distribution of the solutions
found.
In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. Both of them permit to quantify the extent of the Pareto front that is an important factor to consider when evaluating the quality of a Pareto front. Indeed, authors in \cite{zitzler_ec00,weise2012evolutionary} observe that an exploration algorithm should preferably provide Pareto fronts whose objective function range is large. In other words, following the terminology of~\cite{weise2012evolutionary}, a \emph{uniform convergence} should be guaranteed.
Pareto fronts with large objective function ranges permit to investigate in a broader way the potentials of the designed device. This is important, in general, since if only a narrow portion of the Pareto front is provided due to non-uniform convergence, the decision maker may neglect some important options~\cite{weise2012evolutionary}. Having a wide Pareto front becomes essential, in particular, when designing a device that may be required to satisfy different constraints. For example, the decision maker may want to design two different multi-purpose processors. The first design may adapt to situations in which the processor is used for real time elaboration (thus requiring low execution time) and is attached to the powerline (thus not having stringent power dissipation constraints). The second design, on the contrary, may fit situations in which the energy is provided by a battery (stringent power constraints) but only background computations are performed (long execution time is tolerable). These two processors may be successfully designed maintaining the same architecture, using different values for the design parameters. A wide Pareto front helps in achieving this goal.

For a given objective, the \emph{variation range} represents the ratio between the maximum and the minimum value observed for that objective.
\comment{AA: Per rappresentare l'extent del Pereto front, forse sarebbe meglio evitare di usare il variation range, per come l'hai definito tu, e usare la funzione M3 di \cite{zitzler_ec00}, Sec.5 ``Metrics of performance'', che mi pare rappresenti la stessa cosa, concettualmente. Questo ci permette di essere meno attaccabili (del tipo ``non te la prendere con noi, prenditela con Zitzler (che e' citato 3064 volte!) se non capisci l'utilita' della metrica'').}

The average \emph{normalised absolute dispersion error} measures
the average absolute difference between the distribution of points in
the objective space and an ideal distribution in which the points are
uniformly distributed over the objective space. 
Formally, let $O$ be
the image, in the objective space, of the configurations visited by
the design space exploration. The generic element of $O$ (\ie, a
solution) is a pair $(p,t)$ where $p$ and $t$ are the average power
and execution time, respectively. The two-dimensional objective space
is then partitioned by a $M_x \times M_y$ mesh. For each tile $T_i,
\ i=1, 2, \ldots, M_xM_y$ of the mesh, let $N_i$ be the number
of points in $O$ which fall in $T_i$. The average absolute error, $E_i$, for
$T_i$ is the absolute value of the difference between $N_i$ and the
ideal number of solutions, $\overline{N}$, which should fall in $T_i$
in case of uniform distribution. Such $\overline{N}$ can be simply
computed as the ratio between the cardinality of $O$ and the number of
tiles. Thus,
\[ E_i = |N_i - \overline{N}|, \]
where $\overline{N} = |O| / (M_x M_y)$. The average
normalised absolute dispersion error ($ANADE$) is the average of $E_i$
normalised to the maximum absolute error $E_{max}$:
\[ ANADE = \frac{\sum_{i=1}^{M_xM_y} E_i/(M_xM_y)}{E_{max}}, \]
where $E_{max}$ can be computed as the average absolute error in the worst
case in which all the solutions fall in a single tile:
\[ E_{max} = \frac{(M_x M_y - 1) \overline{N} + |\overline{N} -
    |O|| }{M_x M_y}. \]

%\subsection{Impact of the algorithm parameters}
%\secL{parameter_impact}
%\comment{IMPACT OF $K$. Si potrebbero presentare un plot in cui il numero totale di simulazioni e' fissato e vengono eseguiti diversi run dell'algoritmo con diversi valori di K. Per ogni run, il pareto set e' rappresentato nel plot, sottoforma di una curva diversa. Mi aspetterei che tutte le curve siano molto simili. Per ciascun run, il valore di variation range e average absolute dispersion error sono riportati. Mi aspetto valori simili, il che confermerebbe che la diversity nel nostro algoritmo non dipende da $K$. Ne vale la pena rifare questi run? Ci vuole troppo tempo?} 
%
%\comment{Si potrebbe mettere un plot con nelle ascisse l'era e nelle ordinate il numero di split avvenuti in quell'era. Si potrebbero disegnare diverse curve, una per ogni valore di alpha, per far vedere che il numero di split e' piu' alto q. Recommended values are close to 1, as results in \secR{ee} confirm. Using different values, if sufficiently close to 1, does not have a considerable impact on the resulting Pareto front approximation. Ne vale la pena rifare questi run? Ci vuole troppo tempo?}

\subsection{Results}

\figR{pareto_fronts_100} and \figR{pareto_fronts_500} show the
resulting Pareto fronts for two different scenarios of 100 and 500
simulations budget. The set of algorithm parameters involved and their
values are the one previously shown in Table~\ref{tab:defaults}. It
should be pointed out again that the number of generations (or era)
has been chosen so that only the given amount of simulations is
actually performed. Thus, when accounting these simulations only
unique simulations has been considered, removing repetitions and
unfeasible configurations.

% IEEE table style
%\begin{table}
%  \centering
%  \begin{tabular}{ccc}
%    \includegraphics[width=0.30\textwidth]{pictures/alloca_100.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/bmm_100.eps} & 
%    \includegraphics[width=0.30\textwidth]{pictures/fir_int100.eps} \\
%    \includegraphics[width=0.30\textwidth]{pictures/mm_100.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/sha_100.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/wave_100.eps} 
%  \end{tabular}
%  \caption{Pareto fronts found by PS and GA for a fixed budget of 100 configurations.}
%  \label{fig:pareto_fronts_100}
%\end{table}
%
%\begin{table}
%  \centering
%  \begin{tabular}{ccc}
%    \includegraphics[width=0.30\textwidth]{pictures/alloca_500.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/bmm_500.eps} & 
%    \includegraphics[width=0.30\textwidth]{pictures/fir_int500.eps} \\
%    \includegraphics[width=0.30\textwidth]{pictures/mm_500.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/sha_500.eps} &
%    \includegraphics[width=0.30\textwidth]{pictures/wave_500.eps} 
%  \end{tabular}
%  \caption{Pareto fronts found by PS and GA for a fixed budget of 500 configurations.}
%  \label{fig:pareto_fronts_500}
%\end{table}

\begin{figure}
  \figLC{pareto_fronts_100}{Pareto fronts found by PS and GA for a
  fixed budget of 100 configurations.}
  \begin{center}
    \includegraphics[width=0.30\textwidth]{pictures/alloca_100.eps}
    \includegraphics[width=0.30\textwidth]{pictures/bmm_100_bis.eps}
    \includegraphics[width=0.30\textwidth]{pictures/fir_int100.eps}
    \includegraphics[width=0.30\textwidth]{pictures/mm_100.eps}
    \includegraphics[width=0.30\textwidth]{pictures/sha_100_bis.eps}
    \includegraphics[width=0.30\textwidth]{pictures/wave_100.eps} 
  \end{center}
\end{figure}

\begin{figure}
  \figLC{pareto_fronts_500}{Pareto fronts found by PS and GA for a
  fixed budget of 500 configurations.}
  \begin{center}
    \includegraphics[width=0.30\textwidth]{pictures/alloca_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/bmm_500_bis.eps}
    \includegraphics[width=0.30\textwidth]{pictures/fir_int500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/mm_500.eps}
    \includegraphics[width=0.30\textwidth]{pictures/sha_500_bis.eps}
    \includegraphics[width=0.30\textwidth]{pictures/wave_500.eps} 
  \end{center}
\end{figure}

Low budget Pareto sets are shown in \figR{pareto_fronts_100}. Note
that with a so small number of configuration being investigated the
two sets are often overlapping, being also strongly dependent upon the
application being considered.  These results are obtained in about 5
eras and demonstrate that with a so limited budget PS can compete with
genetic approach while even discovering very different Pareto fronts.
Considering the extended simulation budget as depicted in
\figR{pareto_fronts_500}, we can qualitatively observe that even if PS
does not strictly outperforms genetic in terms of dominance, it shows
in some cases largely Pareto in terms of extension. We can intuitively
relate this to the ``novelty-based'' score system, so that instead of
focusing on getting better points inside a given range it tries to
enlarge the range itself, resulting in less dense but more extended
Paretos. Of course we cannot classify this different behaviour as
``better'' or ``worse'' as compared to the genetic one, it depends on
various design factors such as desired results granularity, prefixed
objective range constraints, estimated error of the models used etc.

A comparison between the variation ranges of both power dissipation
and execution time, obtained by PS and GA for different benchmarks is
shown in \figR{dispersion}(a) and (b) respectively.  As it can
be observed, the PS exploration provides solutions which fall on a
range that is, on average, wider than the one provided by a GA
exploration for power dissipation and execution time, respectively.
%\begin{table}
%  \centering
%  \begin{tabular}{c}
%    \includegraphics[width=0.7\textwidth]{pictures/variation_power.eps} \\
%    (a) \\
%    \includegraphics[width=0.7\textwidth]{pictures/variation_etime.eps} \\
%    (b) \\
%    \includegraphics[width=0.7\textwidth]{pictures/dispersion.eps} \\
%    (c) 
%  \end{tabular}
%  \caption{(a)(b) Variation range obtained by PS and GA for different benchmarks
%  and average dispersion absolute dispersion (c)}
%  \label{fig:dispersion}
%\end{table}

\begin{figure}
  \figLC{dispersion}{(a)(b) Variation range obtained by PS and GA for different benchmarks and average dispersion absolute dispersion (c). \comment{AA: manca il titolo nella figura (c)} }
  \begin{center}
    \subfigure[]{\includegraphics[width=0.7\textwidth]{pictures/variation_power.eps} }
    \subfigure[]{\includegraphics[width=0.7\textwidth]{pictures/variation_etime.eps} }
    \subfigure[]{\includegraphics[width=0.7\textwidth]{pictures/dispersion.eps} }
  \end{center}
\end{figure}

\figR{dispersion}(c) shows the average normalised absolute
dispersion errors for different benchmarks for PS and GA
exploration. As it can be observed, GA shows a tendency in
discovering more uniformly dispersed points (i.e. higher values in the
chart). These results, together with the variation ranges of \figR{dispersion}(a) and
(b) confirm the of density vs extension trade-offs
when comparing PS and GA approaches. Again, it
is worth pointing out how a definitive claim about the superiority of
the PS behaviour is strictly dependent on the design 
requirements and constraints, e.g. whether we need wide or dense
results.

\comment{Da lasciare, da spostare ? da mostrare in modo semplice ?  non so}
Another interesting feature of PS is that the final output of its
execution is not only the set of pareto optimal configurations, but also a set
of more or less innovation regions of the parameter space, obtained as
a result of the merge/splitting operators described
in~\secR{algorithm}.

%Nevertheless, the proposed algorithm introduces a new
%perspective on the allocation of a simulations budget, paving the way for
%the development of further strategies for addressing the problem of
%multi-objective hardware/software design space exploration.